---
output:  
    word_document :
        reference_docx: QaTableStyle6.docx
        toc: yes
params: 
    DBServerName: 
        label: "Database Server Name:"
        input: text
        value: ""
        placeholder: "Enter your VDW Database Server Name"
    DBPort: 
        label: "Database Server Port:"
        input: text
        value: ""
        placeholder: "Enter the database port if required.  Leave blank for default."
    DBName:
        label: "Database Name:"
        input: text
        value: ""
        placeholder: "Enter your VDW Database Name"
    DBEncrypt:
        label: "Encrypt Database Connection"
        value: FALSE
    DBUser:
        label: "Database User Name"
        input: text
        value: ""
        placeholder: "leave blank if using windows authentication for your server"
    DBPassword:
        label: "Database Password"
        input: password
        value: ""
        placeholder: "leave blank if using windows authentication for your server"
    QAAlert:
        label: "Create Alert List in Output"
        value: TRUE
---
```{r setup, include=FALSE}
library(RODBC)
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE,
                message = FALSE,
                warning = FALSE,
                error= FALSE)
startTime <- Sys.time()
```

```{r connection strings, include=FALSE, error=FALSE}
connectionString <- getConnectionString(params)
```

```{r run queries, include=FALSE, error=TRUE}
run_query <- function(query_text, ...){
  result <- run_db_query(connectionString, query_text )
  return(result)
}

get_connection <- function(){
  connection <- get_new_connection(connectionString)
  return(connection)
}

```

```{r table replacements, include=FALSE, error=FALSE}
tbresult <-runTableReplacements(connectionString)
```

## CHORDS QA Report: VDW P3 Tables

The purpose of the data quality program is to conduct data quality checks that correspond to similar data checks carried out by PCORnet (Version 7). The checks cover 4 realms of data quality: data model conformance, data plausibility, data completeness, and data persistence. The program uses a series of SQL queries operationalized using RStudio to produce this report. The program is designed to support tables/figures that provide detailed information about each data quality check as well as an alert if a data error appears to be present (based on criteria developed by PCORnet or the CHORDS team). 

This data quality report was generated from CHORDS `r params$DBName`.

## Information about the QA program 
Data Partner: 
Analyst: 
Query Run Date:  `r Sys.Date()`

## Data Check 1.01: Required tables are not present

```{r 1.01 echo = F, include=T, error=T}
# Table to check for quality
tableCheckList <- c(lab_results,   encounters, prescribing,procedures  , social_history  , provider_specialty,diagnoses ,vital_signs  , census_location, demographics  , pro_surveys   ,  pro_questions ,pro_responses)

# table list in DB
sqlConnection <- get_connection()
dbTables <- sqlTables(get_connection())  
RODBC::odbcClose(sqlConnection)

tabledChecked <- vector(length=length(tableCheckList))
for (i in 1:length(tableCheckList)){
  tabledChecked[i] <- sum(grepl(tableCheckList[i], dbTables$TABLE_NAME, ignore.case = T))>0
}

missingTables <- setdiff(tableCheckList, tableCheckList[tabledChecked])
nonMissingTables <- setdiff(tableCheckList, missingTables)
nonMissingDBTables <- subset(dbTables, grepl(paste(tableCheckList, collapse = '|'), TABLE_NAME, ignore.case = T), select=TABLE_NAME)[,,drop=T]

missingTableMessage <- if(length(missingTables)==0) {"All tables accounted for."
} else{
    paste("The following tables are missing:", paste(missingTables, collapse = ','), sep = ' ')
          }
```
 

`r missingTableMessage`


## Data Check 1.02: Required tables are not populated

```{r 1.02 echo = F, include=T, error=T}
table_N_lst <- list() #vector(length=length(nonMissingDBTables))
for(i in 1:length(nonMissingDBTables)){
  table_N_lst[[i]] <- run_query(paste("select distinct '_aaa_' as _aaa_, count(*) as tabN from ",nonMissingDBTables[i], sep=' '))#[,'tabN',drop=T]
 # print(dbTables[i])


  if (class(table_N_lst[[i]]) !='data.frame') {table_N_lst[[i]] <- NULL}
}
#length(dbTables)
table_N_lst <- do.call('rbind', table_N_lst) [,'tabN',drop=T]

nonPopTables <- nonMissingDBTables[table_N_lst==0]
nonPopMessage <- if(length(nonPopTables)==0) {"All tables populated, among those present."
} else{
    paste("The following tables exist, but are not populated:",  paste(nonPopTables, collapse = ','), sep = ' ')
          }
```

`r nonPopMessage`




## Data Check 1.03: Required fields are not present


## Data Check 1.04: Required fields do not conform to data model specifications for data type, length, or name


## Data Check 1.05: Tables have primary key definition errors
```{r 1.05 Data PrimaryKeyChecks Data }
primary_key_data <- run_query("
/*TABLEREPLACEMENTINFILE*/
/*****************************************************************************
Script for validating the VDW Primary Keys.  It will create two lists.  One of
the Expected Tables and their primary keys and another of all the tables currently in
the partner VDW.  It will validate:
    - Compare the expected primary key to the primary key of each
	table in a partner's VDW
	- The number of records that currently violate the exepcted primary key

FOR DEVELOPERS:
To Update the script for partner use:

1) Edit the #PKVALIDATION table insertion if the primary keys of the VDW have
changed or add new entires for new tables.  One row per primary key column.

Examples below:

Output:
    One output table
    1) 	Shows the Table name, the expected primary key, the found primary key,
	and the result of the comparision between the expected and found primary keys.
	KeyMatchResultValues:
		- OK: Expected primary key matches the found primary key
		- KEY MISMATCH: Expected Primary Key does not matc the found primary key
		- TABLE OR KEYS NOT FOUND: No table in the VDW matches the expected table
		- UNKNOWN: An unknow error occured
	ViolationCount:
		- The number of records that that appear to violate the expected primary key.
*****************************************************************************/

/*****************************************************************************
BEGIN TempTable Clearing and Creation
*****************************************************************************/
SET NOCOUNT ON;
BEGIN
   
    IF OBJECT_ID('tempdb..#PKVALIDATION') IS NOT NULL
    BEGIN
	   DROP TABLE #PKVALIDATION;
    END;

	CREATE TABLE #PKVALIDATION (
		Table_Name      VARCHAR(250) ,
        Column_Name     VARCHAR(250)
    );

	IF OBJECT_ID('tempdb..#KeyMatchResult') IS NOT NULL
    BEGIN
	   DROP TABLE #KeyMatchResult;
    END;

	IF OBJECT_ID('tempdb..#PKVIOLATIONS') IS NOT NULL
    BEGIN
	   DROP TABLE #PKVIOLATIONS;
    END;
    
	CREATE TABLE #PKVIOLATIONS (
		Table_Name      VARCHAR(250) ,
        Violation_Count BIGINT
    );

END;
/*****************************************************************************
END TempTable Clearing and Creation
*****************************************************************************/

/*****************************************************************************
BEGIN Table Create Section
*****************************************************************************/

INSERT INTO #PKVALIDATION
VALUES      (
       'CENSUS_DEMOG', 'CENSUS_YEAR'), (
       'CENSUS_DEMOG', 'GEOCODE'), (
       'EVERNDC', 'NDC'), (
       'EVERNDC', 'GENERIC'), (
       'PROVIDER_SPECIALTY', 'PROVIDER'), (
       'DEATH', 'PERSON_ID'), (
       'CAUSE_OF_DEATH', 'PERSON_ID'), (
       'CAUSE_OF_DEATH', 'COD'), (
       'DEMOGRAPHICS', 'PERSON_ID'), (
       'LINKAGE', 'LINK_ID'), (
       'LINKAGE', 'LINE'), (
       'BENEFIT', 'BENEFIT_ID'), (
       'ENCOUNTERS', 'ENC_ID'), (
       'DIAGNOSES', 'DIAGNOSES_ID'), (
       'ENROLLMENT', 'PERSON_ID'), (
       'ENROLLMENT', 'ENR_START'), (
       'LAB_RESULTS', 'LAB_RESULTS_ID'), (
       'PRO_SURVEYS', 'PRO_ID'), (
       'PRO_QUESTIONS', 'PRO_ID'), (
       'PRO_QUESTIONS', 'QUESTION_ID'), (
       'PRO_QUESTIONS', 'QUESTION_VER'), (
       'PRO_RESPONSES', 'RESPONSE_ID'), (
       'PHARMACY', 'PHARMACY_ID'), (
       'PRESCRIBING', 'PRESCRIBING_ID'), (
       'PROCEDURES', 'PROCEDURES_ID'), (
       'SOCIAL_HISTORY', 'SOCIAL_HISTORY_ID'), (
       'VITAL_SIGNS', 'VITAL_SIGNS_ID'), (
       'TUMOR', 'TUMOR_ID'), (
       'LANGUAGES', 'PERSON_ID'), (
       'LANGUAGES', 'LANG_ISO'), (
       'CENSUS_LOCATION', 'PERSON_ID'), (
       'CENSUS_LOCATION', 'LOC_START');
/*****************************************************************************
BEGIN Table Name Replacement: If a TableName replacement table exists, it will swap out the 
names in the tables for the correct ones based on how it's mapped in their table.
*****************************************************************************/

BEGIN
IF OBJECT_ID('CHORDS_TABLENAMES') IS NOT NULL
    BEGIN
        UPDATE a
               SET a.Table_Name = b.NEW_NAME
        FROM #PKVALIDATION a JOIN CHORDS_TABLENAMES b ON b.ORG_NAME = a.Table_Name;
    END;
END;

/*****************************************************************************
END Table Name Replacement
*****************************************************************************/

/*****************************************************************************
BEGIN Analysis Section: Compares the partner's primary keys to the expected keys
******************************************************************************/
SELECT * INTO #KeyMatchResult
FROM (
SELECT            
       ExpectKeys.TABLE_NAME, 
       ExpectKeys.COLUMN_NAMES AS Expected_Primary_Key, 
       CurrKeys.COLUMN_NAMES AS Found_Primary_Key,
       CASE
           WHEN CurrKeys.TABLE_NAME IS NULL
           THEN 'TABLE OR KEYS NOT FOUND'
           WHEN ExpectKeys.COLUMN_NAMES != CurrKeys.COLUMN_NAMES
           THEN 'KEY MISMATCH'
           WHEN ExpectKeys.COLUMN_NAMES = CurrKeys.COLUMN_NAMES
           THEN 'OK'
           ELSE 'UNKNOWN ERROR'
       END AS Key_Match_Result
FROM                  
(
    SELECT        
           TABLE_NAME, 
           LEFT(COL, LEN(COL) - 1) AS COLUMN_NAMES
    FROM              
    (
        SELECT DISTINCT    
               TAB.TABLE_NAME TABLE_NAME, 
        (
            SELECT 
                   COL.Column_Name + ', ' AS [text()]
            FROM    
                 #PKVALIDATION COL
            WHERE  COL.Table_Name = TAB.Table_Name
            ORDER BY 
                     COL.Column_Name FOR
            XML PATH('')
        ) COL
        FROM 
             #PKVALIDATION TAB
    ) T
    WHERE T.COL IS NOT NULL
) ExpectKeys
LEFT JOIN
(
    SELECT        
           TABLE_NAME, 
           LEFT(COL, LEN(COL) - 1) AS COLUMN_NAMES
    FROM              
    (
        SELECT DISTINCT    
               TAB.TABLE_NAME TABLE_NAME, 
        (
            SELECT 
                   COL.COLUMN_NAME + ', ' AS [text()]
            FROM    
                 INFORMATION_SCHEMA.CONSTRAINT_COLUMN_USAGE COL
            WHERE  COL.CONSTRAINT_NAME = TAB.CONSTRAINT_NAME
                   AND COL.TABLE_NAME = TAB.TABLE_NAME
                   AND CONSTRAINT_TYPE = 'PRIMARY KEY'
            ORDER BY 
                     COL.COLUMN_NAME FOR
            XML PATH('')
        ) COL
        FROM 
             INFORMATION_SCHEMA.TABLE_CONSTRAINTS TAB
    ) T
    WHERE T.COL IS NOT NULL
) CurrKeys
     ON CurrKeys.TABLE_NAME = ExpectKeys.TABLE_NAME) KeyMatch;

BEGIN
    
    DECLARE @SQL NVARCHAR(3000);
    DECLARE @Table_Name VARCHAR(100);
	DECLARE @Key_Columns VARCHAR(100);
    
    DECLARE CUR CURSOR
    FOR SELECT
              Table_Name, Expected_Primary_Key
        FROM
             #KeyMatchResult;
    
    OPEN CUR;
    
    FETCH NEXT FROM CUR INTO @Table_Name, @Key_Columns;
    
    WHILE @@FETCH_STATUS = 0
        BEGIN
		SET @SQL = '
			IF OBJECT_ID(''' + @Table_Name + ''') IS NOT NULL
			WITH CTEKEY
				AS (SELECT 
			        COUNT(*) KEYTOT
					FROM (
						SELECT DISTINCT  ' + @Key_Columns + '
						FROM ' + @Table_Name + ') z),
			CTETOT
				AS (SELECT COUNT(*) TABTOT
					FROM ' + @Table_Name + ')
			INSERT INTO #PKVIOLATIONS
			SELECT  ''' + @Table_Name + ''', CTETOT.TABTOT - CTEKEY.KEYTOT
			FROM CTETOT, CTEKEY;'
            --PRINT @SQL;
            EXEC Sp_executesql
                 @SQL;
            FETCH NEXT FROM CUR INTO @Table_Name, @Key_Columns;
        END;
    
    CLOSE CUR;
    
    DEALLOCATE CUR; 
END

SELECT 
       m.*, 
       v.Violation_Count
FROM   
     #KeyMatchResult m
     LEFT JOIN #PKVIOLATIONS v
          ON v.Table_Name = m.TABLE_NAME;
/*****************************************************************************
END Analysis Section
*****************************************************************************/
  ")

```

```{r 1.05 Data PrimaryKeyChecks Result}

knitr::kable(primary_key_data)

```

## Data Check 1.06: Required fields contain values outside of data model specifications


## Data Check 1.07: Required fields have non-permissible missing values


## Data Check 1.08: Tables contain orphan PERSON_IDs 


## Data Check 1.09: Tables contain orphan ENCOUNTER_IDs 


## Data Check 1.10: Replication errors between the ENCOUNTER, PROCEDURES and DIAGNOSIS tables 

```{r 1.10 echo = F, include=T, error=T}

repErrors <- run_query(
                      paste0(
                        "
select 'Diagnosis' as tablename, 
       count(*) as nrows,
       sum(case when a.person_id <> b.person_id then 1 else 0 end) as Person_ID,
       sum(case when a.adate <> b.adate then 1 else 0 end) as Adate,
       sum(case when a.enctype <> b.enctype then 1 else 0 end) as EncType
from ", encounters, " as a inner join (select distinct enc_id, person_id, adate, enctype from ", diagnoses, ") as b on a.enc_id = b.enc_id

union

select 'Procedure' as tablename, 
       count(*) as nrows,
       sum(case when a.person_id <> b.person_id then 1 else 0 end) as  Person,
       sum(case when a.adate <> b.adate then 1 else 0 end) as  Adate,
       sum(case when a.enctype <> b.enctype then 1 else 0 end) as  EncType
from ", encounters, " as a inner join (select distinct enc_id, person_id, adate, enctype from ", procedures, ") as b on a.enc_id = b.enc_id

                        "
                      )) 

repErrors_tx <- tidyr::gather(repErrors, 'field','n_bad',Person_ID, Adate, EncType) %>% arrange(tablename) %>%
  within({
    NP_bad<- paste0(n_bad,' (',round(100*n_bad/nrows, 2),')')

  })

knitr::kable(subset(repErrors_tx, select=c(tablename, nrows, field, NP_bad)),
             col.names = c('Table','N','Field','Errors N(%)'))

```


## Data Check 1.11: More than 5% of encounters are assigned to more than one patient 


```{r 1.11 echo = F, include=T, error=T}

GR1Person <- run_query(
                      paste0(
                        "
select 'Encounter' as tablename, count(*) as n_encid, 
       sum(case when N_person>1 then 1 else 0 end) as N_gr1

from ( select enc_id,
       count(distinct person_id) as N_person
       from ", encounters," 
       group by enc_id
) as qry
"
))

GR1Person <- within(GR1Person, {
  pct <- round(100*N_gr1/n_encid, 2)
})

#if(GR1Person$pct>5){
#  print(paste0("WARNING: The encounter table has more than 5% of encounters assigned to more than 1 person (",GR1Person$pct,"%)"))
#} else{
#  print(paste0("The encounter table has ", GR1Person$pct,"% of encounters assigned to more than 1 person"))
#}

encItoManyMsg <- if(GR1Person$pct>5){
  paste0("WARNING: The encounter table has more than 5% of encounters assigned to more than 1 person (",GR1Person$pct,"%)")
} else{
  paste0("The encounter table has ", GR1Person$pct,"% of encounters assigned to more than 1 person")
}
```


`r encItoManyMsg`


## Data Check 1.12: Tables contain orphan PROVIDER_IDs


## Data Check 1.13: More than 5% of ICD, CPT, LOINC, RXCUI, or NDC codes do not conform to the expected length or content


## Data Check 2.01: More than 5% of records have future dates

```{r 2.01 echo = F, include=T, error=T}
# DC 2.01
tabDates <- run_query('select distinct a.name as tabname, b.name as colname from sys.tables as a inner join sys.columns as b on a.object_id = b.object_id order by a.name') %>% 
  subset(grepl('date',colname, ignore.case = T))
  
modDates <- run_query('select name, modify_date from sys.tables')

futureDateCode <- merge(tabDates, modDates, by.x='tabname', by.y='name') %>%
  within({
    sqlcode <- paste0("select '",tabname, "' as tableName, '", colname,"' as dateName ,count(*) as nrows, sum(case when a.",colname,">b.modify_date then 1 else 0 end) as futureDate from ",
tabname, " as a , (select modify_date from sys.tables where name ='",tabname,"') as b"

)
  })

test <- run_query(futureDateCode[1, "sqlcode"])

fdatelst <- list()
for (i in 1:nrow(futureDateCode)){
  fdatelst[[i]] <- run_query(futureDateCode[i, "sqlcode"])
}

fdatesdone <- do.call('rbind', fdatelst) %>%
  within(.,{
    pctFuture <- round(100*futureDate/nrows, 4)
  })

#params$QAAlert<-F
### add message if no table is a problem 
futureDataMsg <- NULL
futureDataMsg <- if(nrow(subset(fdatesdone, pctFuture>=5))>0 & params$QAAlert==F){
                 knitr::kable(subset(fdatesdone, pctFuture>=5), col.names = c('Table','Date name','Total rows','N future dates','%'))
} else if(nrow(subset(fdatesdone, pctFuture>=5))==0 & params$QAAlert==F) {
                   "No table exceeds the 5% limit on future dates."
} else if ( params$QAAlert==T) {
                   knitr::kable(fdatesdone, col.names = c('Table','Date name','Total rows','N future dates','%'))
                 }
```

`r futureDataMsg`

## Data Check 2.02: More than 10% of records fall into the lowest or highest categories of age, height, weight, diastolic
## blood pressure, systolic blood pressure, or dispensed days supply

```{r 2.02 echo = F, include=T, error=T}
# select count(*) as personCount, sum(case when (minage not between 0 and 89) or (maxage not between 0 and 89) then 1 else 0 end ) as N_outOfRange

patAgeRange <- run_query(paste0("
select count(*) as nrows, 
       sum(case when (minage <0) or (maxage <0) then 1 else 0 end ) as N_low,
       sum(case when (minage >=90) or (maxage >=90) then 1 else 0 end ) as N_high
from
(
select distinct  b.*, datediff(YY, a.birth_date, b.first_dt) as minage,  datediff(YY, a.birth_date, b.last_dt) as maxage 
from ",demographics, " as a inner join 
  (select person_id, min(adate) as first_dt, max(adate) as last_dt from ",encounters, "  group by person_id  ) as b on a.person_id = b.person_id    
where a.birth_date is not null
) as qry
                        "))

htOutOfRange <- run_query(paste0("
                             
select count(*) as nrows,
       sum(case when ht <0.0 then 1 else 0 end ) as N_low,
       sum(case when ht>=95.0 then 1 else 0 end ) as N_high
from ", vital_signs, " 
 where ht is not null
                             "))

wtOutOfRange <- run_query(paste0("
                             
select count(*) as nrows,
       sum(case when wt <0.0 then 1 else 0 end ) as N_low,
       sum(case when wt>350 then 1 else 0 end ) as N_high
from ", vital_signs, " 
 where wt is not null
                             "))


dbpOutOfRange <- run_query(paste0("
                             
select count(*) as nrows,
       sum(case when diastolic <40 then 1 else 0 end ) as N_low,
       sum(case when diastolic>120 then 1 else 0 end ) as N_high
from ", vital_signs, " 
 where diastolic is not null
                             "))
sbpOutOfRange <- run_query(paste0("
                             
select count(*) as nrows,
       sum(case when systolic <40 then 1 else 0 end ) as N_low,
       sum(case when systolic>210 then 1 else 0 end ) as N_high
from ", vital_signs, " 
 where systolic is not null
                             "))

                            
daysSupplyOutOfRange <- run_query(paste0("
                             
select count(*) as nrows,
       sum(case when rx_days_supply <1 then 1 else 0 end ) as N_low,
       sum(case when rx_days_supply>90 then 1 else 0 end ) as N_high
from ", prescribing, " 
 where rx_days_supply is not null
                             "))

allOutOfRange <- 
  rbind(
    cbind(table='Demographic/Encounter', item='Age (people)', low='< 0 yrs',high='> 89 yrs', patAgeRange),
    cbind(table='Vital_signs',item='Height (records)',low= '< 0 inches',high='> 0 inches', htOutOfRange),
    cbind(table='Vital_signs',item='Weight (records)',low= '< 0 pounds',high='> 350 pounds', wtOutOfRange),
    cbind(table='Vital_signs',item='Diastolic BP (records)',low= '< 40 mgHg',high='> 120 mgHg', dbpOutOfRange),
    cbind(table='Vital_signs',item='Systolic BP (records)',low= '< 40 mgHg',high='> 210 mgHg', sbpOutOfRange),
    cbind(table='Prescribing',item='Prescribed days supply (records)',low= '< 1 day',high='> 90 days', daysSupplyOutOfRange)
)
allOutOfRange <- within(allOutOfRange, {
  NP_low = paste0(N_low,' (',round(100*N_low/nrows, 2),')')
  NP_high = paste0(N_high,' (',round(100*N_high/nrows, 2),')')
})
#knitr::kable(
#subset(allOutOfRange, select=c(table,item, low, high, nrows, NP_low, NP_high)),
#col.names = c('Table','Field','Check Low','High','N','Low values N (%)','High values N (%)')
#)

 
outOfRangeMsg <- NULL
outOfRangeMsg <- if(nrow(subset(allOutOfRange, (N_low+N_high)/nrows>=0.1))>0 & params$QAAlert==F){
   knitr::kable(
   subset(subset(allOutOfRange, (N_low+N_high)/nrows>=0.1), select=c(table,item, low, high, nrows, NP_low, NP_high)),
   col.names = c('Table','Field','Check Low','High','N','Low values N (%)','High values N (%)')
)
} else if(nrow(subset(allOutOfRange, (N_low+N_high)/nrows>=0.1))==0 & params$QAAlert==F) {
                   "No table exceeds the 5% limit."
} else if ( params$QAAlert==T) {
  knitr::kable(
  subset(allOutOfRange, select=c(table,item, low, high, nrows, NP_low, NP_high)),
  col.names = c('Table','Field','Check Low','High','N','Low values N (%)','High values N (%)')
)
                 }

```

`r outOfRangeMsg`



## Data Check 2.03: More than 5% of patients have illogical date relationships

```{r 2.03 echo = F, include=T, error=T}

encTabBirth2.03 <- run_query(paste0("
 select sum(dobPostAdate) as dobPostAdate,
         sum(dobPostDdate) as dobPostDdate
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, b.adate)<0 and b.adate is not null then 1 else 0 end) as dobPostAdate,
                              max(case when datediff(day, a.birth_Date, b.ddate)<0 and b.ddate is not null then 1 else 0 end) as dobPostDdate
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", encounters, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


encTabDeath2.03 <- run_query(paste0("
 select sum(deathPreAdate) as deathPreAdate,
         sum(deathPreDdate) as deathPreDdate
  from (
                      select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.adate)>0 and b.adate is not null then 1 else 0 end) as deathPreAdate,
                              max(case when datediff(day, a.deathdt, b.ddate)>0 and b.ddate is not null then 1 else 0 end) as deathPreDdate
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", encounters, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


procTabBirth2.03 <- run_query(paste0("
 select sum(dobPostProc) as dobPostProc 
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, procdate)<0 and b.procdate is not null then 1 else 0 end) as dobPostProc
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", procedures, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


procTabDeath2.03 <- run_query(paste0("
 select sum(deathPreProc) as deathPreProc
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.procdate)>0 and b.procdate is not null then 1 else 0 end) as deathPreProc
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", procedures, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )
vitalsTabBirth2.03 <- run_query(paste0("
 select sum(dobPostMeasure) as dobPostMeasure 
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, b.measure_date)<0 and b.measure_date is not null then 1 else 0 end) as dobPostMeasure
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", vital_signs, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


vitalsTabDeath2.03 <- run_query(paste0("
 select sum(deathPreMeasure) as deathPreMeasure
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.measure_date)>0 and b.measure_date is not null then 1 else 0 end) as deathPreMeasure
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", vital_signs, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )
pharmTabBirth2.03 <- run_query(paste0("
 select sum(dobPostRx) as dobPostRx 
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, b.rxdate)<0 and b.rxdate is not null then 1 else 0 end) as dobPostRx
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", pharmacy, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


pharmTabDeath2.03 <- run_query(paste0("
 select sum(deathPreRx) as deathPreRx
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.rxdate)>0 and b.rxdate is not null then 1 else 0 end) as deathPreRx
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", pharmacy, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )
  
prescribTabBirth2.03 <- run_query(paste0("
 select sum(dobPostRxStart) as dobPostRxStart 
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, b.rx_start_date)<0 and b.rx_start_date is not null then 1 else 0 end) as dobPostRxStart
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", prescribing, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


prescribTabDeath2.03 <- run_query(paste0("
 select sum(deathPreRxStart) as deathPreRxStart
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.rx_start_date)>0 and b.rx_start_date is not null then 1 else 0 end) as deathPreRxStart
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", prescribing , " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )

labTabBirth2.03 <- run_query(paste0("
 select sum(dobPostResult) as dobPostResult 
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.birth_Date, b.result_dt)<0 and b.result_dt is not null then 1 else 0 end) as dobPostResult
                       from (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as a inner join
                            ", lab_results, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


labTabDeath2.03 <- run_query(paste0("
 select sum(deathPreResult) as deathPreResult
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.result_dt)>0 and b.result_dt is not null then 1 else 0 end) as deathPreResult
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                            ", lab_results, " as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )


DeathBeforeBirth <- run_query(paste0("
 select sum(deathPreBirth) as deathPreBirth
  from (
                       select a.person_id, 
                              max(case when datediff(day, a.deathdt, b.birth_date)>0  then 1 else 0 end) as deathPreBirth
                       from (select distinct person_id, deathdt from ",death, " where deathdt is not null) as a inner join
                             (select distinct person_id, birth_date from ",demographics, " where birth_date is not null) as b on a.person_id = b.person_id group by a.person_id

) qry

                       
                       
                       ") )

patsInEncTab <- run_query( paste0("select count(distinct person_id) as n from ",encounters))

table2.03 <- cbind(
encTabBirth2.03,
encTabDeath2.03,
procTabBirth2.03,
procTabDeath2.03,
vitalsTabBirth2.03,
vitalsTabDeath2.03,
pharmTabBirth2.03,
pharmTabDeath2.03,
prescribTabBirth2.03,
prescribTabDeath2.03,
labTabBirth2.03,
labTabDeath2.03,
DeathBeforeBirth
)

labels2.03 <- data.frame(name=c( "dobPostAdate","dobPostDdate","deathPreAdate","deathPreDdate","dobPostProc","deathPreProc","dobPostMeasure","deathPreMeasure","dobPostRx","deathPreRx","dobPostRxStart","deathPreRxStart","dobPostResult","deathPreResult","deathPreBirth" ),
                        newlabel = c('Adate<birth_date','Ddate<birth_date','Deathdt<adate','Deathdt<ddate','Procdate<birth_date','ProcDate>deathdt','Measure_date<birth_date','Measure_date>deathdt','rxdate<birth_date','rxdate>deathdt','rx_start_dt<birth_date','rx_start_dt>deathdt','Result_date<birth_date','Result_date>deathdt','Deathdt<birth_date')   ,
                        srctab = c('Encounters','Encounters','Encounters','Encounters','Procedures','Procedures','Vital_signs','Vital_signs','Pharmacy','Pharmacy','Prescribing','Prescribing','lab_results','lab_results','Demographics and death')
)
 
table2.03_tx <- tidyr::gather(table2.03, "comp", 'n') %>%
  within({
    pct <- round(100*n/patsInEncTab$n, 2)
  }) 
table2.03_tx$ord <- 1:nrow(table2.03_tx)
table2.03_tx <- table2.03_tx%>%
  merge(labels2.03, by.x='comp',by.y='name', all.x=T) %>%
  arrange(ord)

knitr::kable(table2.03_tx[,c('ord','comp','n','pct', 'srctab')], col.names = c(' ','Date comparison','N Patients','% of unique patients in the encounter table','Source table'))

#params$QAAlert <- T
illogicalDatesMsg <- NULL
illogicalDatesMsg <- if(nrow(subset(table2.03_tx, pct>=5))>0 & params$QAAlert==F){
        knitr::kable(subset(table2.03_tx, pct>=5)[,c('ord','comp','n','pct', 'srctab')], 
                    col.names = c(' ','Date comparison','N Patients','% of unique patients in the encounter table','Source table'))
} else if(nrow(subset(fdatesdone, pctFuture>=5))==0 & params$QAAlert==F) {
                   "No table exceeds the 5% limit on illogical dates."
} else if ( params$QAAlert==T) {
        knitr::kable(table2.03_tx[,c('ord','comp','n','pct', 'srctab')], 
                    col.names = c(' ','Date comparison','N Patients','% of unique patients in the encounter table','Source table'))
                 }

```

`r illogicalDatesMsg`



## Data Check 3.03: More than 10% of records have missing or unknown values for the following fields:

```{r 3.03 echo=F}

# to implement a new check add a new row to the following table:
# table = the standard name of the table in the database (it will automaticaly correct for you if the actual table names are not standard)
# var = the name of the variable to be chacked
# condition = a valid logical condition ( in MS sql server language) which identifies a bad value
# where = a valid where clause to subset the table, be sure to start with "WHERE", see example for ddate
check3.03 <-
rbind( 
  data.frame(table='demographics',   var='birth_date',            condition='is null', where=' '),
  data.frame(table='demographics',   var='gender',                condition="not in ('M','F','O')", where=' '),
  data.frame(table='encounters',     var='discharge_disposition', condition="not in ('A','E')", where=' '),
  data.frame(table='encounters',     var='ddate',                 condition="is null", where="where enctype = 'IP'"),
  data.frame(table='procedures',     var='procdate',              condition='is null', where=' '),
  data.frame(table='prescribing',    var='rx_order_date',         condition='is null', where=' '),
  data.frame(table='pharmacy',       var='not(rxsup>1 or rxamt>1)',       condition=' ', where=' '), # compicated by it depending on two variables
  #data.frame(table='cause_of_death', var='source',                condition="not in ('S', 'N', 'T', 'B', 'L', 'U', 'O')"),
  data.frame(table='diagnoses',      var='dx_origin',             condition="not in ('OD', 'BI', 'CL', 'PR', 'NI', 'OT')", where=' '),
  data.frame(table='diagnoses',      var='enc_id',                condition="is null", where=' '),
  data.frame(table='procedures',     var='enc_id',                condition="is null", where=' '),
  data.frame(table='vital_signs',    var='enc_id',                condition="is null", where=' ')
  
)

res3.03 <- list()
for(i in 1:nrow(check3.03)){
  
  
  res3.03[[i]] <- with(check3.03, {run_query( 
                                            paste("select '",table[i],"' as dataTable,'",var[i],"' as variable, count(*) as N_rows, sum(case when ",var[i],condition[i],"then 1 else 0 end) as N_bad",
                                                  "from",eval(as.name(tolower(table[i]))), where[i]
                                                  ,
                                                  sep=' '
                                                  )
  )}
  )
}
result3.03 <- do.call('rbind', res3.03) %>%
  within(.,{
    pctBad <- round(100*N_bad/N_rows, 2)
  })

check3.03Msg <- NULL
check3.03Msg <- if(nrow(subset(result3.03, pctBad>=10))>0 & params$QAAlert==F){
        knitr::kable(subset(result3.03, pctBad>=10)[,c('dataTable','variable','N_rows','N_bad', 'pctBad')], 
                    col.names = c('Table','Variable','N Rows','N with bad values','%'))
} else if(nrow(subset(fdatesdone, pctFuture>=5))==0 & params$QAAlert==F) {
                   "No check exceeds the 10% limit on bad values."
} else if ( params$QAAlert==T) {
        knitr::kable(result3.03[,c('dataTable','variable','N_rows','N_bad', 'pctBad')], 
                    col.names = c('Table','Variable','N Rows','N with bad values','%'))
                 }


```

`r check3.03Msg`



## Data Check 3.04: Less than 50% of patients with encounters have DIAGNOSIS records 


## Data Check 3.05: Less than 50% of patients with encounters have PROCEDURES records 
```{r 3.05 echo = F, include=T, error=T}

patsWithEncs <- run_query(paste0("select count(*) as nrows, sum(patsWEncDiag) as patsWEncDiag, sum(patsWEncProc) as patsWEncProc from ( select  case when person_id in (select person_id from ", diagnoses, ") then 1 else 0 end as patsWEncDiag",
 ",case when person_id in (select person_id from ", procedures, ") then 1 else 0 end as patsWEncProc",
 "       from ", demographics, " where person_id in (select person_id from ", encounters, ") ) as q") )

patsWithEncs <- within(patsWithEncs, {
  pctWEncDiag <- 100* patsWEncDiag/nrows
  pctWEncproc <- 100* patsWEncProc/nrows
})

patDiagMsg <- with (patsWithEncs,{
  if(pctWEncDiag<50 ){
  paste0('WARNING: Only ',round(pctWEncDiag,1),'% of patients with encounters have diagnoses') 
} else { 
  paste0('NOTE: ',round(pctWEncDiag,1),'% of patients with encounters have diagnoses') 
}
})

patProcMsg <- with (patsWithEncs,{
  if(pctWEncproc<50 ){
  paste0('WARNING: Only ',round(pctWEncproc,1),'% of patients with encounters have procedures') 
} else { 
  paste0('NOTE: ',round(pctWEncproc,1),'% of patients with encounters have procedures') 
}
})

knitr::kable(
  rbind(
    cbind(check='DC 3.05', result=patDiagMsg),
    cbind(check='DC 3.04', result=patProcMsg)
  )
)
```


## Data Check 3.06: More than 10% of IP (inpatient) or ED to inpatient (EI) encounters with any diagnosis don't have a principal
## diagnosis


## Data Check 3.07: Encounters, diagnoses, or procedures in an ambulatory (AV), emergency department (ED), ED to inpatient
## (EI), or inpatient (IP) setting are less than 75% complete three months prior to the current month
```{r 3.07 Data, include=FALSE, error=FALSE}

benchmark_start <- run_query(paste0("
  SET NOCOUNT ON;
  DECLARE @BenchmarkStartDate DATE;
  SET @BenchmarkStartDate =
  (
      SELECT 
             IIF(MAX(x.ADATE) <= GETDATE(), MAX(x.ADATE), GETDATE())
      FROM ",   
           encounters, " x
  );
  SELECT @BenchmarkStartDate;
"))

data_result <- run_query(str_replace_all(paste0("
  SET NOCOUNT ON;
  DECLARE @BenchmarkPriorYearAvg INT;
  SET @BenchmarkPriorYearAvg =
  (
      SELECT 
             COUNT(1) / 12
      FROM  ", 
           encounters, " e
      WHERE  e.ENCTYPE IN('IP', 'AV', 'ED', 'EI')
      AND e.ADATE >= DATEADD(MONTH, -24, @BenchmarkStartDate)
      AND e.ADATE <= DATEADD(MONTH, -12, @BenchmarkStartDate)
  );
  
  SELECT 
         FORMAT(ADATE, 'MM-MMM') MONTHNUM, 
         @BenchmarkPriorYearAvg AS BenchMarkCount, 
         COUNT(1) PriorMonthCount, 
         ROUND(CAST(COUNT(1) AS FLOAT) / @BenchmarkPriorYearAvg * 100.0, 2) AS PercentOfBenchMark
  FROM ",  
       encounters, " e
  WHERE  e.ENCTYPE IN('IP', 'AV', 'ED', 'EI')
  AND e.ADATE >= DATEADD(MONTH, -12, @BenchmarkStartDate)
  GROUP BY 
           FORMAT(ADATE, 'MM-MMM')
  ORDER BY 
           FORMAT(ADATE, 'MM-MMM');                          
                            
"), "@BenchmarkStartDate", paste0('\'', benchmark_start[1,1], '\'', sep='')))
```

```{r 3.07 Output, error=FALSE, include=TRUE}

knitr::kable(data_result)

```

## Data Check 3.11: Vital, prescribing, or laboratory records are less than 75% complete three months prior to the current month


####Total program run time:
```{r calc_runtime, echo=TRUE}
# close odbc
odbcCloseAll()
endTime <- Sys.time()
runtime <- endTime - startTime
```


Query run time = `r runtime` minutes
